---
title: "Model Assessment"
author: "Juan Martínez"
date: "12/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
library(knitr)
library(dplyr)
library(tidyr)
library(STRbook)
library(tidyverse)
library(caret)
library(knitcitations)
library(knitLatex)
library(ISLR)
library(boot)
library(tibble)
library(wesanderson)
library(kableExtra)
library(stringr)
library(dplyr)
library(ggplot2)
library(likelihood)
library(anytime)
library(lubridate)
library(tibble)
library(data.table)
library(splines)
library(mlbench)
library(ggfortify)
library(RColorBrewer)
library(leaflet)
library(cowplot)
library(compare)
library(pscl)
library(MASS)
library(glmmTMB)
library(mgcv)
library(lmtest)

```

# Introduction

We start by importing the data set. The file source is `cartaapc_dashboard.csv`. Then, we convert the character variables into factor variables:

```{r hists, echo = T, fig.width=10, fig.height=8}

carta <- read_csv("Data/Chattanooga/cartaapc_dashboard.csv")

head(carta)
```


Given the size of the `carta` `DateFrame`, we divide the data in two halves:

- `carta1`: `direction_id == 1`
- `carta0`: `direction_id == 0`

After that, the main `DateFrame` is removed (or erased) to have more space available:

```{r vartime, echo = T, eval = F}

carta1 <- carta %>%
  filter(direction_id == 1) %>%
  mutate(Hour = lubridate::hour(arrival_time),
         Day = factor(wday(date),
                      levels = c("1","2", "3", "4", "5","6" , "7"),
                      labels = c("Sunday","Monday","Tuesday", 
                                 "Wednesday","Thursday", "Friday",
                                 "Saturday"))) %>%
  summarize(trip_id, arrival_time, stop_id, stop_sequence, 
            stop_lat, stop_lon, route_id, board_count,
            alight_count, occupancy, date, service_period,
            trip_start_time, Day)
  
head(carta1)
write.csv(carta1, "carta1.csv", row.names = F)

rm(carta)
```

## Data Partition Training Data

The data has a hierchical order:

- routes
  * trips
  
Thus, we need to get information keeping this in mind:

```{r carta1, echo = T}

carta1_routes <- unique(carta1$route_id)

route1_4 <- carta1 %>%
  filter(route_id == carta1_routes[10]) %>%
  summarize(trip_id, arrival_time, stop_id, stop_sequence, 
            stop_lat, stop_lon, route_id, board_count,
            alight_count, occupancy, date, trip_start_time, Day)

rm(carta1)
head(route1_4)

route1_4$trip_id <- factor(route1_4$trip_id)
route1_4$route_id <- factor(route1_4$route_id)

carta1_trips_id_route_4 <- route1_4 %>%
  summarise(Uniq_trips = unique(trip_id))

carta1_stop_id_route_4 <- route1_4 %>%
  summarise(Uniq_stops = unique(stop_id))


```


The following function (`data_for_reg_board()`) extracts and converts the number of board counts (`board_count`) and hour of the day (`Hour_f`) per bus stop (`stop_id`) per route `route_id`:


```{r SunDatTrain, echo = T, fig.width=10, fig.height=8}

#----------------------------------------------------------------------
# stop_id:
#----------------------------------------------------------------------

data_for_reg_occupancy <- function(Da, st, Dat){
  Data_reg <- Dat %>%
    filter(Day %in% Da, stop_id == st) %>%
    mutate(Hour = hour(arrival_time),
           Hour_f = as.factor(Hour)) %>%
    group_by(Hour) %>%
    summarise(occupancy, Hour_f, trip_id)
  
   Data_reg <- Data_reg[ , -c(1)]
   
   neg_occup <- which((Data_reg$occupancy < 0) == T, arr.ind = T)
  
   Data_reg$occupancy[neg_occup] = 0

  return(Data_reg)
}

zero_hour_check <- function(dat){
  zero_hour <- which((dat$Hour_f == 0) == T, arr.ind = T)
  Data_reg <- dat[-c(zero_hour),]
  return(Data_reg)
}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
weekends <- c("Saturday", "Sunday")

route1_4_stop_1919_week <- data_for_reg_occupancy(weekdays,
                                                  carta1_stop_id_route_4$Uniq_stops[25],
                                                  route1_4)

#----------------------------------------------------------------------
# Data partition: 70% and 30%

train_prop = 0.7
#----------------------------------------------------------------------
# Training Data: 
#----------------------------------------------------------------------


get_set <- function(prop, Dat, set){
  set.seed(123)
  len = dim(Dat)[1]
  train_ind = sample(len, ceiling(len*prop))
  
  if(set == "train"){
    data_train = Dat[train_ind, ]
    return(data_train)
  }
  else{
    data_test = Dat[-train_ind, ]
    return(data_test)
  }
}


#----------------------------------------------------------------------
# Training Data: 
#----------------------------------------------------------------------

route1_4_stop_1919_week_train <- get_set(0.7, route1_4_stop_1919_week, "train")

#----------------------------------------------------------------------
# Testing Data: 
#----------------------------------------------------------------------

route1_4_stop_1919_week_test <- get_set(0.7, route1_4_stop_1919_week, "test")
#----------------------------------------------------------------------

```

# Overall Occupancy:

- Day of the week
-
```{r}


route1_4_stop_1919_week_train %>%
  group_by(Hour_f) %>%
  mutate(lambda = mean(occupancy)) %>%
  ggplot(aes(x = Hour_f, y = occupancy)) +
  geom_point(alpha = 0.01, color = 'red') +
  geom_point(aes(x = Hour_f, y = lambda),
             alpha = 0.6, color = "red") +
  xlab("Hour of the day") +
  ylab("Occupancy") +
  ggtitle("Overall Occupancy for Route 4:")


```
# Generalized Linear Models (Poisson):

From the Training data set, we build GLM for every day of the week considering a Poisson distribution.

## Weekdays GLM Poisson

### Stop 1919:

- Take into account that predictions need to be given for each trip! 
- Get the average rate for each bus stop-trip.

```{r SundGLMPois, echo = T}
#------------------------------------------------------------------------------
# Stop 1919: 

K=10


k_fold_cv_Pois <- function(k, dat, dat_test){
  part = floor(dim(dat)[1]/k)
  sm <- list()
  MSE = 0
  k_fold_pred <- list()
  predictions = 0
  set.seed(123)
  rand_ind <- matrix(sample(part*k, part*k, replace = F),
                     nrow = k)
  ind = 0
  model_aic = 0
  cv.error = rep(0, k)
  for(i in 1:k){
    glm.fit = glm(occupancy ~ Hour_f,
                       family = poisson(link = "log"),
                       data = dat[-rand_ind[i, ], ])
    sm[[i]] = glm.fit
    k_fold_pred[[i]] = predict(glm.fit, dat[rand_ind[i, ], ],
                               type = "response")
    MSE[i] = mean((dat$occupancy[rand_ind[i, ]] - k_fold_pred[[i]])^{2})
  }
  ind = which(MSE == min(MSE), arr.ind = T)
  train_MSE = MSE[ind]
  model = sm[[ind]]
  test <- dat_test %>%
    summarise(occupancy, Hour_f)
  predictions = predict(model, test, type = "response")
  test_MSE = mean((test$occupancy - predictions)^{2})
  performance <- list("MSE" = MSE, "train_MSE" = train_MSE,
                      "Best_Model" = model,
                      "test_MSE" = test_MSE,
                      "Predicted" = predictions)
  return(performance)
}

#------------------------------------------------------------------------------

Weekdays_Pois_cv <- k_fold_cv_Pois(10, route1_4_stop_1919_week_train,
                                   route1_4_stop_1919_week_test)

route1_4_stop_1919_week_test$Predictions <- Weekdays_Pois_cv[["Predicted"]]

route1_4_stop_1919_week_test_Pois <- route1_4_stop_1919_week_test %>%
  group_by(Hour_f) %>%
  mutate(lambda = mean(occupancy))
  

route1_4_stop_1919_week_test_Pois %>%
  group_by(Hour_f) %>%
  ggplot(aes(x = Hour_f, y = occupancy)) +
  geom_point(alpha = 0.01, color = "red") +
  geom_point(aes(x = Hour_f, y = lambda),
             alpha = 0.6, color = "red") +
  geom_point(aes(x = Hour_f, y = Predictions),
             alpha = 0.5, color = "blue")

```

### Stop 1883:

```{r}

route1_4_stop_1883_week <- data_for_reg_occupancy(weekdays,
                                                  carta1_stop_id_route_4$Uniq_stops[1],
                                                  route1_4)

#----------------------------------------------------------------------
# Data partition: 70% and 30%

train_prop = 0.7
#----------------------------------------------------------------------
# Training Data: 
#----------------------------------------------------------------------

route1_4_stop_1883_week_train <- get_set(0.7, route1_4_stop_1883_week, "train")
route1_4_stop_1883_week_test <- get_set(0.7, route1_4_stop_1883_week, "test")
#----------------------------------------------------------------------

K = 10

Weekdays_1883_Pois_cv <- k_fold_cv_Pois(K, route1_4_stop_1883_week_train,
                                        route1_4_stop_1883_week_test)

route1_4_stop_1883_week_test$Predictions <- Weekdays_1883_Pois_cv[["Predicted"]]

route1_4_stop_1883_week_test_Pois <- route1_4_stop_1883_week_test %>%
  group_by(Hour_f) %>%
  mutate(lambda = mean(occupancy))
  

route1_4_stop_1883_week_test_Pois %>%
  group_by(Hour_f) %>%
  ggplot(aes(x = Hour_f, y = occupancy)) +
  geom_point(alpha = 0.01, color = "red") +
  geom_point(aes(x = Hour_f, y = lambda),
             alpha = 0.6, color = "red") +
  geom_point(aes(x = Hour_f, y = Predictions),
             alpha = 0.5, color = "blue")


```


### Stop 923:

```{r}
route1_4_stop_923_week <- data_for_reg_occupancy(weekdays,
                                                  carta1_stop_id_route_4$Uniq_stops[2],
                                                  route1_4)

#----------------------------------------------------------------------
# Data partition: 70% and 30%

train_prop = 0.7
#----------------------------------------------------------------------
# Training Data: 
#----------------------------------------------------------------------

route1_4_stop_923_week_train <- get_set(0.7, route1_4_stop_923_week, "train")
route1_4_stop_923_week_test <- get_set(0.7, route1_4_stop_923_week, "test")
#----------------------------------------------------------------------

K = 10

Weekdays_923_Pois_cv <- k_fold_cv_Pois(K, route1_4_stop_923_week_train,
                                       route1_4_stop_923_week_test)

route1_4_stop_923_week_test$Predictions <- Weekdays_923_Pois_cv[["Predicted"]]

route1_4_stop_923_week_test_Pois <- route1_4_stop_923_week_test %>%
  group_by(Hour_f) %>%
  mutate(lambda = mean(occupancy))
  

route1_4_stop_923_week_test_Pois %>%
  group_by(Hour_f) %>%
  ggplot(aes(x = Hour_f, y = occupancy)) +
  geom_point(alpha = 0.01, color = "red") +
  geom_point(aes(x = Hour_f, y = lambda),
             alpha = 0.6, color = "red") +
  geom_point(aes(x = Hour_f, y = Predictions),
             alpha = 0.5, color = "blue")

```

### Stop 956:

```{r}
route1_4_stop_956_week <- data_for_reg_occupancy(weekdays,
                                                  carta1_stop_id_route_4$Uniq_stops[35],
                                                  route1_4)

#----------------------------------------------------------------------
# Data partition: 70% and 30%

train_prop = 0.7
#----------------------------------------------------------------------
# Training Data: 
#----------------------------------------------------------------------

route1_4_stop_956_week_train <- get_set(train_prop, route1_4_stop_956_week, "train")
route1_4_stop_956_week_test <- get_set(train_prop, route1_4_stop_956_week, "test")
#----------------------------------------------------------------------

K = 10

Weekdays_956_Pois_cv <- k_fold_cv_Pois(K, route1_4_stop_956_week_train,
                                       route1_4_stop_956_week_test)

route1_4_stop_956_week_test$Predictions <- Weekdays_956_Pois_cv[["Predicted"]]

route1_4_stop_956_week_test_Pois <- route1_4_stop_956_week_test %>%
  group_by(Hour_f) %>%
  mutate(lambda = mean(occupancy))
  

route1_4_stop_956_week_test_Pois %>%
  group_by(Hour_f) %>%
  ggplot(aes(x = Hour_f, y = occupancy)) +
  geom_point(alpha = 0.01, color = "red") +
  geom_point(aes(x = Hour_f, y = lambda),
             alpha = 0.6, color = "red") +
  geom_point(aes(x = Hour_f, y = Predictions),
             alpha = 0.5, color = "blue")

```


### Stop 957:

```{r}


route1_4_stop_957_week <- data_for_reg_occupancy(weekdays, carta1_stop_id_route_4$Uniq_stops[35], route1_4)
route1_4_stop_957_week <-zero_hour_check(route1_4_stop_957_week)

#----------------------------------------------------------------------
# Data partition: 70% and 30%

train_prop = 0.7
#----------------------------------------------------------------------
# Training Data: 
#----------------------------------------------------------------------

route1_4_stop_957_week_train <- get_set(train_prop, route1_4_stop_957_week, "train")
route1_4_stop_957_week_test <- get_set(train_prop, route1_4_stop_957_week, "test")
#----------------------------------------------------------------------

K = 10

Weekdays_957_Pois_cv <- k_fold_cv_Pois(K, route1_4_stop_957_week_train,
                                       route1_4_stop_957_week_test)
sort(unique(route1_4_stop_957_week_train$Hour_f))
route1_4_stop_957_week_test$Predictions <- Weekdays_957_Pois_cv[["Predicted"]]

route1_4_stop_957_week_test_Pois <- route1_4_stop_957_week_test %>%
  group_by(Hour_f) %>%
  mutate(lambda = mean(occupancy))
  

route1_4_stop_957_week_test_Pois %>%
  group_by(Hour_f) %>%
  ggplot(aes(x = Hour_f, y = occupancy)) +
  geom_point(alpha = 0.01, color = "red") +
  geom_point(aes(x = Hour_f, y = lambda),
             alpha = 0.6, color = "red") +
  geom_point(aes(x = Hour_f, y = Predictions),
             alpha = 0.5, color = "blue")

```


Let's make a for lop:

```{r}

K = 10
train_prop = 0.7
Route4_dir1_stops_week_Pois_cv <- list()
Route4_dir1_stops_week_test <- list()

for(i in 1:34){
  data <- data_for_reg_occupancy(weekdays, carta1_stop_id_route_4$Uniq_stops[i], route1_4)
  data_train <- get_set(train_prop, data , "train")
  data_test <- get_set(train_prop, data , "test")
  Route4_dir1_stops_week_Pois_cv[[i]] <- k_fold_cv_Pois(K, data_train, data_test)            
  Route4_dir1_stops_week_test[[i]][["Predictions"]] <- Route4_dir1_stops_week_Pois_cv[[i]][["Predicted"]]
}


Route4_dir1_stops_week_Pois_cv[[35]] <- Weekdays_957_Pois_cv
Route4_dir1_stops_week_test[[35]][["Predictions"]] <- Weekdays_957_Pois_cv[["Predicted"]]

```


## Mondays GLM Poisson


# Appendix {-}

```{r appendix1, echo = T, eval = F}
data_for_reg <- function(Da, st, Y, Dat){
  
  if(Y == "board_count"){
     Data_reg <- Dat %>%
       filter(Day == Da, stop_id == st) %>%
       mutate(Hour = lubridate::hour(Time),
              Hour_f = as.factor(Hour)) %>%
       group_by(Hour) %>%
       summarise(board_count, Hour_f)
  
     Data_reg <- Data_reg[ , -c(1)]
  }
  else if(Y == "alight_count"){
    Data_reg <- Dat %>%
       filter(Day == Da, stop_id == st) %>%
       mutate(Hour = lubridate::hour(Time),
              Hour_f = as.factor(Hour)) %>%
       group_by(Hour) %>%
       summarise(alight_count, Hour_f)
  
    Data_reg <- Data_reg[ , -c(1)]
  }
  else if(Y == "occupancy"){
    Data_reg <- Dat %>%
       filter(Day == Da, stop_id == st) %>%
       mutate(Hour = lubridate::hour(Time),
              Hour_f = as.factor(Hour)) %>%
       group_by(Hour) %>%
       summarise(occupancy, Hour_f)
  
    Data_reg <- Data_reg[ , -c(1)]
  }
  else if(Y == "Load"){
    Data_reg <- Dat %>%
       filter(Day == Da, stop_id == st) %>%
       mutate(Hour = lubridate::hour(Time),
              Hour_f = as.factor(Hour)) %>%
       group_by(Hour) %>%
       mutate(load = board_count - alight_count) %>%
       summarise(load, Hour_f)
  
    Data_reg <- Data_reg[ , -c(1)]
  }
 
  return(Data_reg)
  
}

#-------------------------------------------------------------

#----------------------------------------------------------------------

Stop_id1_1_train <- Data_train %>%
  filter(stop_id == selected_bus_id_1[1]) %>%
  mutate(Hour = lubridate::hour(Time)) %>%
  group_by(Day, Hour) %>%
  summarise(board_count, Date, Day)


Stop_id1_2_train <- Data_train %>%
  filter(stop_id == selected_bus_id_1[2]) %>%
  mutate(Hour = lubridate::hour(Time)) %>%
  group_by(Week_day = as.factor(wday(Date)), Hour) %>%
  summarise(lambda_Hour = mean(board_count), board_count,  Date, Day)


Stop_id1_3_train <- Data_train %>%
  filter(stop_id == selected_bus_id_1[3]) %>%
  mutate(Hour = lubridate::hour(Time)) %>%
  group_by(Week_day = as.factor(wday(Date)), Hour) %>%
  summarise(lambda_Hour = mean(board_count), board_count, Date, Day)


Stop_id1_4_train <- Data_train %>%
  filter(stop_id == selected_bus_id_1[4]) %>%
  mutate(Hour = lubridate::hour(Time)) %>%
  group_by(Week_day = as.factor(wday(Date)), Hour) %>%
  summarise(lambda_Hour = mean(board_count), board_count, Date, Day)



#----------------------------------------------------------------------------------
# Stop 1:
#----------------------------------------------------------------------------------

Sundays1_train <- data_for_reg("Sunday", selected_bus_id_1[1],
                               "board_count", Data_train)

Mondays1_train <- data_for_reg("Monday", selected_bus_id_1[1],
                               "board_count", Data_train)

Tuesdays1_train <- data_for_reg("Tuesday", selected_bus_id_1[1],
                                "board_count", Data_train)

Wednesdays1_train <- data_for_reg("Wednesday", selected_bus_id_1[1],
                                "board_count", Data_train)

Thursdays1_train <- data_for_reg("Thursday", selected_bus_id_1[1],
                                "board_count", Data_train)

Fridays1_train <- data_for_reg("Friday", selected_bus_id_1[1],
                                "board_count", Data_train)
#----------------------------------------------------------------------------------
# Stop 2:
#----------------------------------------------------------------------------------


Sundays2_train <- data_for_reg("Sunday", selected_bus_id_1[2],
                               "board_count", Data_train)

Mondays2_train <- data_for_reg("Monday", selected_bus_id_1[2],
                               "board_count", Data_train)

Tuesdays2_train <- data_for_reg("Tuesday", selected_bus_id_1[2],
                                "board_count", Data_train)

Wednesdays2_train <- data_for_reg("Wednesday", selected_bus_id_1[2],
                                "board_count", Data_train)

Thursdays2_train <- data_for_reg("Thursday", selected_bus_id_1[2],
                                "board_count", Data_train)

Fridays2_train <- data_for_reg("Friday", selected_bus_id_1[2],
                                "board_count", Data_train)

#----------------------------------------------------------------------------------
# Stop 3:
#----------------------------------------------------------------------------------

Sundays3_train <- data_for_reg("Sunday", selected_bus_id_1[3],
                               "board_count", Data_train)

Mondays3_train <- data_for_reg("Monday", selected_bus_id_1[3],
                               "board_count", Data_train)

Tuesdays3_train <- data_for_reg("Tuesday", selected_bus_id_1[3],
                                "board_count", Data_train)

Wednesdays3_train <- data_for_reg("Wednesday", selected_bus_id_1[3],
                                "board_count", Data_train)

Thursdays3_train <- data_for_reg("Thursday", selected_bus_id_1[3],
                                "board_count", Data_train)

Fridays3_train <- data_for_reg("Friday", selected_bus_id_1[3],
                                "board_count", Data_train)

#----------------------------------------------------------------------------------
# Stop 4:
#----------------------------------------------------------------------------------

Sundays4_train <- data_for_reg("Sunday", selected_bus_id_1[4],
                               "board_count", Data_train)

Mondays4_train <- data_for_reg("Monday", selected_bus_id_1[4],
                               "board_count", Data_train)

Tuesdays4_train <- data_for_reg("Tuesday", selected_bus_id_1[4],
                                "board_count", Data_train)

Wednesdays4_train <- data_for_reg("Wednesday", selected_bus_id_1[4],
                                "board_count", Data_train)

Thursdays4_train <- data_for_reg("Thursday", selected_bus_id_1[4],
                                "board_count", Data_train)

Fridays4_train <- data_for_reg("Friday", selected_bus_id_1[4],
                                "board_count", Data_train)


#----------------------------------------------------------------------------



#------------------------------------------------------------------------------
# Stop 1: 

Sun_stop2.mod_Hour <- glm(board_count ~ Hour_f, family =  poisson(link = "log"),
                          data = Sundays2_train)

Sun_stop2.mod_Hour.loglike <- logLik(Sun_stop2.mod_Hour)[1]

#------------------------------------------------------------------------------
# Stop 1: 

Sun_stop3.mod_Hour <- glm(board_count ~ Hour_f, family = poisson(link = "log"),
                          data = Sundays3_train)

Sun_stop3.mod_Hour.loglike <- logLik(Sun_stop3.mod_Hour)[1]

#------------------------------------------------------------------------------
# Stop 1: 

Weekdays_1919_Pois <- glm(occupancy ~ Hour_f, family = poisson(link = "log"),
                          data = Weekdays_stop_1919_train)

Sun_stop4.mod_Hour.loglike <- logLik(Sun_stop4.mod_Hour)[1]

#------------------------------------------------------------------------------
# Performance data: 
#------------------------------------------------------------------------------

#------------------------------------------------------------------------------
# AIC:

Sun_stop1.mod_Hour_AIC <- AIC(Sun_stop1.mod_Hour)
Sun_stop2.mod_Hour_AIC <- AIC(Sun_stop2.mod_Hour)
Sun_stop3.mod_Hour_AIC <- AIC(Sun_stop3.mod_Hour)
Sun_stop4.mod_Hour_AIC <- AIC(Sun_stop4.mod_Hour)

Sundays_train_glm_AIC <- data.frame(rbind(c(Sun_stop1.mod_Hour_AIC,
                                            Sun_stop2.mod_Hour_AIC,
                                            Sun_stop3.mod_Hour_AIC,
                                            Sun_stop4.mod_Hour_AIC)))

#------------------------------------------------------------------------------
# Log-like:


Sundays_train_glm.loglike <- data.frame(rbind(c(Sun_stop1.mod_Hour.loglike,
                                            Sun_stop2.mod_Hour.loglike,
                                            Sun_stop3.mod_Hour.loglike,
                                            Sun_stop4.mod_Hour.loglike)))

colnames(Sundays_train_glm_AIC) <- paste("stop_id_",
                                         selected_bus_id_1,
                                         sep = "")
rownames(Sundays_train_glm_AIC) <- c("Poisson")
Sundays_train_glm_AIC




```


