---
title: "Occupancy  Data Analysis"
author: "Juan Martínez"
date: "12/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
library(knitr)
library(dplyr)
library(tidyr)
library(STRbook)
library(tidyverse)
library(caret)
library(knitcitations)
library(htmltools)
library(htmlwidgets)
library(knitLatex)
library(sparkline)
library(tibble)
library(plotly)
library(wesanderson)
library(kableExtra)
library(stringr)
library(dplyr)
library(ggplot2)
library(likelihood)
library(ggmap)
library(ggnetwork)
library(anytime)
library(lubridate)
library(tibble)
library(data.table)
library(rgdal)
library(raster)
library(sp)
library(splines)
library(mlbench)
library(ggfortify)
library(RColorBrewer)
library(sf)
library(maps)
library(leaflet)
library(cowplot)
library(compare)
library(pscl)
library(MASS)
library(glmmTMB)
library(mgcv)
library(lmtest)
library(readr)
```

# Introduction

In this document we run some analysis about occupancy. We group the information by `route_id`, `stop_id`, `date`, and hour of the day. Then, we calculate the median, $75^{th}$ and $90^{th}$ percentiles of the "grouped" occupancies. These values are considered as thresholds that help us identify trips with anomalous occupancy values.

```{r datainput, echo = T}

Weekdays_Occupancy <- read_csv("Data/Chattanooga/Weekdays_Occupancy_Analysis_by_trip_id.csv")

head(Weekdays_Occupancy)

```

# Filtering by `direction_id`

## `direction_id == 1`

```{r}

carta1 <- Weekdays_Occupancy %>%
  filter(direction_id == 1)

```

### Median Occupancy

Finding the trips whose occupancy is higher than the median occupancy per route per date:

```{r}
route_median1 <- carta1 %>%
  group_by(route_id, stop_id, date = date(date_time), hour = hour(date_time)) %>%
  mutate(route_median_occup = median(occupancy)) %>%
  filter(occupancy > route_median_occup) %>%
  summarise(route_id, trip_id, date, hour, stop_id, occupancy, route_median_occup)

head(route_median1)

```


Exploratory data analysis:

```{r}
route_median1_hour <- route_median1 %>%
  group_by(route_id, stop_id, trip_id, hour) %>%
  mutate(z_score = (occupancy - mean(occupancy))/sd(occupancy)) %>%
  summarise(occupancy, mean_occupancy = mean(occupancy), z_score)


head(route_median1_hour)
```


### $75^{th}$ Percentile


```{r}
route_percent75_1 <- carta1 %>%
  group_by(route_id, stop_id, date = date(date_time), hour = hour(date_time)) %>%
  mutate(route_percent75_occup = quantile(occupancy, prob = 0.75)) %>%
  filter(occupancy > route_percent75_occup) %>%
  summarise(route_id, trip_id, hour, stop_id, occupancy, route_percent75_occup)

head(route_percent75_1)
```

Z-scores:

```{r}
route_percent75_1_hour <- route_percent75_1 %>%
  group_by(route_id, stop_id, trip_id, hour) %>%
  mutate(z_score = (occupancy - mean(occupancy))/sd(occupancy)) %>%
  summarise(occupancy, mean_occupancy = mean(occupancy), z_score)


head(route_percent75_1_hour)
```



### $90^{th}$ Percentile

## `direction_id == 0`

```{r}

carta0 <- Weekdays_Occupancy %>%
  filter(direction_id == 0)

```


### Median Occupancy

Finding the trips whose occupancy is higher than the median occupancy per route per date:

```{r}
route_median0 <- carta0 %>%
  group_by(route_id, stop_id, date = date(date_time), hour = hour(date_time)) %>%
  mutate(route_median_occup = median(occupancy)) %>%
  filter(occupancy > route_median_occup) %>%
  summarise(route_id, trip_id, date, hour, stop_id, occupancy, route_median_occup)

head(route_median0)

```

# Descriptive data analysis


```{r desc1, echo = T}

Weekdays_median_anomaly <- Weekdays_Occupancy %>%
  summarise(trip_id, stop_id, route_id, occupancy, date_time, Mean_trip, Median_trip)

head(Weekdays_median_anomaly)

```

## ti 

```{r desc2, echo = T}
# rm(Weekdays_Occupancy)

Weekdays_median_anomaly_stop_date <- Weekdays_median_anomaly %>%
  mutate(date = date(date_time),
         hour = hour(date_time)) %>%
  group_by(route_id, stop_id, date, hour) %>%
  filter(occupancy > Median_trip) %>%
  summarise(occupancy, Median_trip, stop_id)

head(Weekdays_median_anomaly_stop_date)
```

